---
title: "Introduction to R"
author: "W. Joel Schneider"
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r}
library(tidyverse)
library(psych)
library(rio)
library(broom)
```

# Import Data

```{r}
d <- import("https://github.com/wjschne/RWorkshop/raw/master/Data.csv")
```

# Descriptive Statistics

```{r d, echo = TRUE}
#overall Summary
summary(d)
# More complete descriptives (from psych package)
describe(d)
# Describe by groups (from psych package)
describeBy(d,group = "Treatment")
# Mean of a single variable
mean(d$RC1)
# Standard deviation
sd(d$RC1)

```

# Analysis

## Does reading comprehension at time 1 predict reading comprehension at time 2?

```{r}
m1 <- lm(RC2 ~ RC1, data = d)
summary(m1)
```

## After controlling for reading comprehension, is treatment associated with higher reading comprehension at time 2?

```{r}
m2 <- lm(RC2 ~ RC1 + Treatment, data = d)
summary(m2)
```

## Is the treatment effect stronger for students with initially lower reading comprehension?

```{r}
m3 <- lm(RC2 ~ RC1 * Treatment, data = d)
summary(m3)
```

# Getting Started With Plots

We use data.frame `d` and tell which variables are "mapped" onto which scales. Here, `RC1` is on the x-axis, and `RC2` is on the y-axis.

```{r}
ggplot(data = d, mapping = aes(x = RC1, y = RC2)) 
# Shorter code
ggplot(d, aes(RC1, RC2))
```

But the plot is empty! Let's add a new layer onto the plot using `geom_point`.

```{r}
ggplot(d, aes(RC1, RC2)) + 
  geom_point()  
```

Not bad, but the plot does not make our point yet. We want to show the effect of treatment on the time 2 reading comprehension scores. Let's color the points by the treatment group variable.

```{r}
ggplot(d, aes(RC1, RC2, color = Treatment)) + 
  geom_point()
```

Better, but it is still not easy to see what is going on. Let's add regression lines.

```{r}
ggplot(d, aes(RC1, RC2, color = Treatment)) + 
  geom_point() + 
  geom_smooth()
```

Cool! But those are loess regression lines, not the straight regression lines we are used to. We can use the `lm` (linear model) method:

```{r}
ggplot(d, aes(RC1, RC2, color = Treatment)) + 
  geom_point() + 
  geom_smooth(method = "lm") 
```

Okay, let's extend both lines all the way across the plot:

```{r}
ggplot(d, aes(RC1, RC2, color = Treatment)) + 
  geom_point() + 
  geom_smooth(method = "lm", fullrange = TRUE) 
```

Add some ellipses to make the point crystal clear:

```{r}
ggplot(d, aes(RC1, RC2, color = Treatment)) + 
  geom_point() + 
  geom_smooth(method = "lm", fullrange = TRUE) + 
  stat_ellipse()
```

